---
title: "Central Limit Theorem"
subtitle: ""
abstract: "This is a presentation about Central Limit Theorem"
keywords: "central limit theorem, R"
titlegraphic: "../images/itu.eps"
titlegraphic-width: 0.15
author:
  - \href{mailto:sezenismail@gmail.com}{İ.SEZEN}
  - \href{mailto:alper.unal@gmail.com}{A.ÜNAL}
institute:
  - Eurasia Institute of Earth Sciences
institute-short: İTÜ
date: \today
fontsize: 10pt
navigation: horizontal
graphics: yes
graphicspath:
  - \graphicspath{{LLN_files/figure-beamer/}}
beamerconfig:
    blocks: # default rounded, no shadow. Do not recommend shadows.
      shadow:
section-titles: # Hide section title pages
colorlinks: yes
linkcolor: "crimson"
handout: true
toc: true
toc-title: Content
toc-depth: 3
justifying: yes
bibliography: ../bibliography.bib
link-citations: true
output:
  beamer_presentation:
    template: ../template/default-2.9.2.1.beamer
    keep_tex: yes
    slide_level: 3
    fig_caption: yes
    theme: CambridgeUS
    colortheme: euresia
    highlight: tango
    pandoc_args: ["--pdf-engine-opt=-shell-escape", "--listings", "-H", "../template/listings-setup.tex"]
    includes:
      in_header: ../template/before_body.tex
      after_body: ../template/after_body.tex
    
---

```{r setup, include=FALSE}
source("../setup.R")
n <- 10

plot_dist <- function(dat, mu = NULL, sd = NULL) {
  mu_y_pos <- diff(range(dat$y))
  p <- ggplot(dat, aes_string("x", "y")) + 
    geom_line(col = "orange", size = 1) +
    labs(x = "Quantiles", y = "Density")

  if (!is.null(mu)) {
    p <- p + geom_vline(xintercept = mu, linetype = "dashed", col = "black") +
      annotate("text", x = mu, y = mu_y_pos/10, hjust = -0.5,
               label = TeX("$\\mu$", output = "character"), parse = TRUE)
  }
  if (!is.null(sd)) {
    p <- p + 
      geom_segment(aes(x = mu, y = 0, xend = mu - sd, yend = 0),
                   arrow = arrow(length = unit(0.05, "inches"), 
                                 ends = "both")) +
      annotate("text", x = mu - sd, y = mu_y_pos/30, hjust = -0.5,
               label = TeX("$\\sigma$", output = "character"), parse = TRUE)
  }
  return(p)
}



# https://www.researchgate.net/post/How_does_one_generate_random_values_of_my_own_function_distribution
ddist <- function(x) {
  (dnorm(x, 40, 10) + dnorm(x, 60, 5) + dnorm(x, 90, 15) + 
     dnorm(x, 150, 10))/4
}

pdist <- function(q) {
    (pnorm(q, 40, 10) + pnorm(q, 60, 5) + pnorm(q, 90, 15) + 
     pnorm(q, 150, 10))/4
}

qdist <- function(p) {
  sapply(p, function(p) {
    if (p < 0 | p > 1) return(NaN)
    if (p == 0) return(-Inf)
    if (p == 1) return(Inf)
    uniroot(function(q, p) p - sapply(q, pdist), c(-1e3, 1e3), p = p,
            tol = sqrt(.Machine$double.eps))$root
  })
}

rdist <- function(n) qdist(runif(n))

evdist <- function() {
  integrate(function(x) x * ddist(x), -Inf, Inf)$value
}

sddist <- function() {
  sqrt(integrate(function(x) x^2 * ddist(x), -Inf, Inf)$value - evdist()^2)
}
```

# Definition
##
### Definition {.fragile}
<!-- .allowframebreaks -->

_Central Limit Theorem_ states that if you have a population with mean $\mu$ and standard deviation $\sigma$ and take sufficiently large random samples from a population regardless of the distribution with replacement, then the distribution of the sample means will be approximately normally distributed^[[1]]^.

That is, a histogram of the sample means of a large number of different batches ($k$) of the same kind of data, each of size $n$, will look like a bell curve (_Gaussian sampling distribution_) if $k$ is large enough [@Wilks2011].

- Independent of the distribution of source population
- The independence of the observations is not really necessary
- The resulting distribution is _always_ Gaussian.
- Standard deviation of means is _always_ less than $\sigma$.

[1]: http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Probability/BS704_Probability12.html


# Showcase
##
### Sampling from Distribution {.fragile}

Assume that you have a population which has a PDF as below;

```{r fig1, echo=FALSE, fig.width = 6, out.width = '90%'}
y <- seq(0, 200, length.out = 200)
dat <- data.frame(x = y, y = ddist(y))
plot_dist(dat)
```

### take $n$ sample from population $k$ times {.fragile}

And take $n = `r n`$ sample from population $15$ times.

```{r fig2, echo=FALSE, fig.width = 7, fig.height = 4, out.width = '90%'}
x <- as.data.frame(sapply(1:15, function(i) {
  set.seed(i); rdist(n)
}))
colnames(x) <- 1:ncol(x)
m <- colMeans(x); pm <- pretty(m)
m <- do.call(rbind, lapply(1:length(m), function(i) {
  x <- m[1:i]
  data.frame(col = 1:length(x), Sample = i, x, y = 0)
}))
m <- cbind(type = factor("Mean"), m)
x <- reshape2::melt(x, id.vars = 0, variable.name = "Sample",
                    value.name = "x")
x <- cbind(type = factor("Sample"), col = x$Sample, x, y = 0)
x <- rbind(x, m)
px <- pretty(x$x)
segs1 <- data.frame(type = factor("Sample"), x = px, xend = px, 
                    y = -0.05, yend = 0.05)
segs2 <- data.frame(type = factor("Mean"), x = pm, xend = pm,
                    y = -0.05, yend = 0.05)
segs <- rbind(segs1, segs2)
ggplot(x, aes(x, y, fill = col)) + labs(x = "") +
  geom_hline(yintercept = 0, size = 0.3) + ylim(c(-0.1, 0.1)) +
  geom_segment(aes(x, y, xend = xend, yend = yend), data = segs,
               size = 0.3, inherit.aes = FALSE) +
  geom_point(aes(shape = type), size = 2, show.legend = FALSE) +
  scale_shape_manual(values = c(21, 24)) +
  facet_grid(Sample ~ type, scales = "free_x", switch = "y") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.y.left = element_text(angle = 0))
```

### Effect of $k$ {.fragile .allowframebreaks}

Take $n = `r n`$ sample from population $k = 15, 100, 500, 1000$ times.

```{r fig3, echo=FALSE, fig.width = 7, fig.height = 4, out.width = '90%'}
k <- c(15, 100, 500, 1000)
x <- as.data.frame(sapply(1:max(k), function(i) {
  set.seed(i); rdist(n)
}))
colnames(x) <- 1:ncol(x)
m <- colMeans(x); pm <- pretty(m)
m <- data.frame(Sample = 1:length(m), x = m)
m <- do.call(rbind, lapply(k, function(k) {
  cbind(k = factor(k), m[m$Sample <= k,])
}))
m <- data.frame(type = factor("Mean"), m)
y <- reshape2::melt(x, id.vars = 0, variable.name = "Sample",
                    value.name = "x")
y$Sample <- as.integer(y$Sample)
y <- do.call(rbind, lapply(k, function(k) {
  cbind(type = factor("Sample"), k = factor(k), y[y$Sample <= k,])
}))
dat <- rbind(y, m)

label_k <- paste0("k=", k); names(label_k) <- k
ploth <- function(dat) {
  ggplot(dat, aes(x)) + labs(x = "", y = "Density") +
  geom_histogram(aes(y=..density..), binwidth = 2, col = "black",
                 fill = "transparent") +
  geom_density(alpha = .2, fill = "#FF6666") +
  facet_grid(k ~ type, scales = "free_y",
             labeller = labeller(k = label_k)) +
      theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank())
}
ploth(dat[as.integer(dat$k) <= 2,])
ploth(dat[as.integer(dat$k) >= 3,])
```

### Effect of $n$ (Law of Large Numbers) {.fragile}

Take $n = 10,100$ sample from population $k = 100$ times.

```{r fig4, echo=FALSE, fig.width = 7, fig.height = 4, out.width = '90%'}
n <- c(10, 100)
x <- as.data.frame(sapply(1:100, function(i) { set.seed(i); rdist(max(n)) }))
colnames(x) <- 1:ncol(x)
xl <- list(x1 = x[1:min(n),], x2 = x)

dat <- do.call(rbind, lapply(xl, function(x) {
  m <- colMeans(x)
  m <- data.frame(type = factor("Mean"), n = nrow(x), Sample = 1:length(m), x = m)
  y <- reshape2::melt(x, id.vars = 0, variable.name = "Sample",
                    value.name = "x")
  y$Sample <- as.integer(y$Sample)
  y <- cbind(type = factor("Sample"), n = nrow(x), y)
  dat <- rbind(y, m)
  dat$n <- factor(dat$n)
  dat
}))
rownames(dat) <- NULL
dat <- dat[dat$x < 200,]

ci <- aggregate(x ~ type + n, dat, function(i) {
  quantile(i, c(0.025, 0.975))
})
ci <- cbind(ci[,1:2], ci[,3])
colnames(ci)[3:4] <- c("c1", "c2")

ci2 <- ci
ci2[ci2$type == "Sample",3:4] <- ci2[ci2$type == "Mean",3:4]

label_n <- paste0("n=", n); names(label_n) <- n
ploth <- function(dat) {
  ggplot(dat, aes(x)) + labs(x = "", y = "Density") +
    geom_density(fill = "transparent") +
    geom_segment(aes(x = c1, xend = c2), y = 0, yend = 0, data = ci, 
                 col = "red", size = 1) +
    geom_rect(aes(xmin = c1, xmax = c2), data = ci2, ymin = -Inf, 
              ymax = Inf, alpha = 0.2, fill = "forestgreen", 
              inherit.aes = FALSE) +
  facet_grid(n ~ type, scales = "free_y",
             labeller = labeller(n = label_n)) +
      theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank())
  }
ploth(dat)
```

# Conclusion
## 
### Conclusion

- If number of samples $k \rightarrow \infty$, distribution of the sample means approximates to _Gauissan distribution_.
- If sample size $n \rightarrow \infty$, standard deviation of the sample means (Standard Error) $SE \rightarrow 0$. 

# References
