---
title: "Hypothesis Testing"
subtitle: "(t-test)"
abstract: "This is a presentation about Hypothesis Testing"
keywords: "hypothesis testing, t-test, R"
---

```{r setup, include=FALSE}
source("setup.R")
```

# \hypt
## What is \hypt?
### What is \hypt?

\hyp\ is,

- a supposition or proposed explanation made on the basis of limited evidence 
  as a starting point for further investigation.
- a testable statement about the relationship **between two or more variables**.
- a proposed explanation for some observed phenomenon^[[1]]^.
- an assumption about a population parameter. This assumption may or may not be 
  true (For instance, a population mean is equal to $10$)^[[2]]^.

[1]: https://whatis.techtarget.com/definition/hypothesis
[2]: https://stattrek.com/statistics/dictionary.aspx?definition=statistical_hypothesis

### What is \hypt?

\hyp\ is,

- a statistical method that is used in making statistical decisions using 
  experimental data^[[3]]^.
- basically an assumption that we make about the population parameter^[[3]]^.
- is a method of [statistical inference](https://en.wikipedia.org/wiki/Statistical_inference)^[[4]]^.
- The theory, methods, and practice of testing a \hyp\ by comparing it 
  with the **\hn{}**.

[3]: https://www.statisticssolutions.com/hypothesis-testing/
[4]: https://en.wikipedia.org/wiki/Statistical_hypothesis_testing

### Usage Cases

- A _sample_ is compared against a population parameter.
- Two statistical data sets are compared^[[3]]^.
- a data set obtained by sampling is compared against a synthetic data set 
  from an idealized model^[[3]]^.
  
### Steps of \hyp\ test

1. Identify a _\ts{}_ that is appropriate to the data and question.
2. Define a **\hn{}**, $H_{0}$.
3. Define an **\ha{}**, $H_{A}$.
4. Obtain the _null distribution_, which is simply the sampling distribution
  for the test statistic, if the _\hn{}_ is _true_.
5. Compare the observed _\ts{}_ to the _null distribution_.

### t-test

1. $H_{0}$: Mean is equal to $\mu$ $\mu - \bar{x} = 0$ ($\pval < \alpha$)
2. $H_{A}$: Mean is different than $\mu$ $\mu - \bar{x} \neq 0$ ($\pval \ge \alpha$)

### Confidence and Alpha Levels {.fragile .allowframebreaks}

1. **Significance ($\alpha$) level:** The probability of rejecting the $H_{0}$ when the null hypothesis is true.
2. **Confidence Level:** The probability of confidence interval contains the true population mean ($\mu$). \textcolor{violet}{It is not the probability of which the $\mu$ lies within interval.}

```{r mycode1, echo = TRUE, caption="T-test", firstnumber=1}
alpha <- 0.05
conf.level <- 1 - alpha
```
