---
title: "Linear Regression - II"
subtitle: "Examples"
abstract: "This is a presentation about linear regression"
keywords: "linear regression, R"
toc-multicol: false
---

```{r setup, include=FALSE}
source("setup.R")
library(forecast)
library(lmtest)
library(ggfortify)
nr <- sum(complete.cases(airquality))
contains <- function(fit, pattern) {
  f <- as.character(fit$call$formula)
  (length(grep(pattern, f))) != 0
}
mylog <- function(fit) {
  fun <- function(x) x
  if (contains(fit, "log")) fun <- function(x) exp(x)
  if (contains(fit, "log10")) fun <- function(x) 10^x
  return(fun)
}
get_res <- function(fit) {
  lg <- mylog(fit)
  r <- resid(fit)
  if (contains(fit, "log")) {
    o <- lg(fit$model[, 1])
    r <- o - o / lg(r)
  }
  return(r)
}
res_std_err <- function(fit) {
  sqrt(sum(get_res(fit)^2) / fit$df.residual)
}
get_metrics <- function(fit) {
  lg <- mylog(fit)
  o <- lg(fit$model[, 1])
  p <- lg(predict(fit))
  rmse <- round(Metrics::rmse(o, p), 2)
  mae <- round(Metrics::mae(o, p), 2)
  res_se <- round(res_std_err(fit), 2)
  return(c(RMSE = rmse, MAE = mae, Res.SE = res_se))
}
```

# airquality Dataset
## Exploratory Analysis
### EDA {.fragile}
<!-- .allowframebreaks -->

Let's explore `airquality` dataset. Rows contains any `NA` were removed for convenience. We have $`r nr`$ obeservation after removal.

```{r}
aq <- airquality
aq$Temp <- ((aq$Temp - 32) * (5 / 9))
aq <- aq[complete.cases(aq), ]
print(head(aq), digits = 0)
```

### Summary {.fragile}

```{r}
summary(aq)
```

### Pairs plot {.fragile}

```{r, echo = FALSE, fig.width = 10, fig.height = 6.5, out.width = '90%'}
nrm <- function(x, min = 0, max = 1, range_x = range(x)) {
  a <- (max - min) / (max(range_x) - min(range_x))
  b <- max - a * max(range_x)
  a * x + b
}
panel <- function(x, y, ...) {
  ok <- is.finite(x) & is.finite(y)
  if (any(ok)) {
    xx <- x[ok]; yy <- y[ok]
    if (par("xlog")) xx <- log10(xx)
    if (par("ylog")) yy <- log10(yy)
    fit <- lm(yy ~ xx)
    abline(fit, col = "Navy")
  }
  panel.smooth(x, y, ...)
}
upper.panel <- function(x, y, ...) {
  u <- par("usr")
  xl <- par("xlog")
  yl <- par("ylog")
  points(x, y, ...)
  panel.smooth(x, y, ...)
  panel(x, y, ...)
  xx <- if (xl) log10(x) else x
  yy <- if (yl) log10(y) else y
  r <- cor(xx, yy, use = "p")
  xx <- nrm(0, u[1], u[2], c(0, 1))
  yy <- nrm(0.9, u[3], u[4], c(0, 1))
  if (par("xlog")) xx <- 10^xx
  if (par("ylog")) yy <- 10^yy
  text(xx, yy, sprintf("r = %.2f", r), cex = 2, adj = -0.1, col = "blue")
}
diag.panel <- function(x, ...) {
  ylog <- par("ylog"); on.exit(par(ylog))
  xlog <- par("xlog"); u <- par("usr")
  par(ylog = FALSE)
  tryd <- try(d <- density(if (xlog) log10(x) else x, na.rm = TRUE),
              silent = TRUE)
  if (class(tryd) != "try-error") {
    if (xlog) d$x <- 10^d$x
    d$y <- nrm(d$y, u[3], u[4] * 0.8)
    lines(d$x, d$y, col = "dodgerblue4", xlog = xlog)
  }
  rug(x)
  axis(1, pretty(x), tcl = 0.5, tck = 0.04, lwd = 0.5, labels = FALSE,
       xlog = xlog)
}
pairs(aq[1:4], panel = panel,
      upper.panel = upper.panel, diag.panel = diag.panel,
      cex = 0.7, pch = 20, gap = 0)
```

### Pairs plot - log of Ozone {.fragile}

```{r, echo = FALSE, fig.width = 10, fig.height = 6.5, out.width = '90%'}
pairs(Ozone ~ Solar.R + Wind + Temp, aq,
      panel = panel, log = 1,
      upper.panel = upper.panel, diag.panel = diag.panel,
      cex = 0.7, pch = 20, gap = 0)
```

## Fit Multiple Linear Regression
### Fitting for Ozone {.fragile .allowframebreaks}

Fit a _multiple linear model_ to predict _Ozone_ concentration.
```{r}
fit1 <- lm(Ozone ~ Solar.R + Wind + Temp, aq)
summary(fit1)
```

### Diagnostics {.fragile .allowframebreaks}

```{r, include = TRUE, fig.width = 10, fig.height = 5, out.width='90%'}
library(ggfortify)
autoplot(fit1)
```

```{r, include = TRUE, fig.width = 7, fig.height = 3, out.width='90%'}
autoplot(fit1, which = c(4, 6))
```

### Prediction for Ozone {.fragile}

```{r, echo = FALSE, results = 'asis'}
p <- data.frame(n = seq_len(nrow(aq)), Original = aq$Ozone,
                fit1 = predict(fit1))
dat <- reshape2::melt(p, id.vars = 1, variable.name = "Type",
                      value.name = "Ozone")
dfe <- data.frame(fit1 = get_metrics(fit1))
print.df(t(dfe))
```

```{r, echo = FALSE, fig.width = 7, fig.height = 4, out.width='90%'}
ggplot(dat, aes(n, Ozone, col = Type)) +
  geom_point() + geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed")
```

### Fitting for log of Ozone {.fragile .allowframebreaks}

Fit a _multiple linear model_ to predict _log of Ozone_ concentration.

```{r}
fit2 <- lm(log10(Ozone) ~ Solar.R + Wind + Temp, aq)
summary(fit2)
```

### Diagnostics {.fragile .allowframebreaks}

```{r, include = TRUE, fig.width = 10, fig.height = 5, out.width='90%'}
autoplot(fit2)
```

```{r, include = TRUE, fig.width = 7, fig.height = 3, out.width='90%'}
autoplot(fit2, which = c(4, 6))
```

### Prediction {.fragile}

```{r, echo = FALSE, results = 'asis'}
p <- cbind(p, fit2 = 10^predict(fit2))
dat <- reshape2::melt(p, id.vars = 1, variable.name = "Type",
                      value.name = "Ozone")
dfe <- cbind(dfe, fit2 = get_metrics(fit2))
print.df(t(dfe))
```


```{r, echo = FALSE, fig.width = 7, fig.height = 3.5, out.width='90%'}
ggplot(dat, aes(n, Ozone, col = Type)) +
  geom_point(size = 0.5) + geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed")
```



## Durbin Watson Test
### Definition {.fragile}

_Durbin–Watson test_ is a test statistic used to detect the presence of autocorrelation at $lag = 1$ in the residuals (prediction errors) from a regression analysis.

```{r, eval = FALSE}
library(lmtest)
dwtest(formula, order.by = NULL,
       alternative = c("greater", "two.sided", "less"),
       iterations = 15, exact = NULL, tol = 1e-10, data = list())
```

* $H_0$ is that there is no correlation among residuals, i.e., they are independent.
* $H_a$ is that residuals are autocorrelated. 

### Interpretation {.fragile .allowframebreaks}

How to interpret _Durbin-Watson test_ output?

```{r}
(dwt1 <- dwtest(fit1))
```

You **fail to reject** $H_0$ at $\alpha = 0.05$ for `fit1`.

\newpage

How to interpret _Durbin-Watson test_ output?

```{r}
(dwt2 <- dwtest(fit2))
```

You **fail to reject** $H_0$ at $\alpha = 0.05$ for `fit2`.
 
### Check Residuals {.fragile .allowframebreaks}

<!--
BG-test: null-hypo is that there is no serial correlation of any order up to n
-->

```{r, fig.width = 8, fig.height = 2.5, out.width = '90%'}
forecast::checkresiduals(fit1)
```

\newpage

```{r, fig.width = 8, fig.height = 2.5, out.width = '90%'}
forecast::checkresiduals(fit2)
```

## Principal Component Analysis
### How to use PCA for linear Regression {.fragile .allowframebreaks}

We can use _PCA_ to set up a linear model. _R_ has two functions:

* `prcomp`: Spectral Decomposition
* `princomp`: Singular Value Decomposition

```{r}
(pca <- prcomp(aq[, 2:4], center = TRUE, scale. = TRUE))
```

<!-- pca2 <- princomp(aq[,2:4], cor = TRUE) -->

\newpage

```{r}
(spca <- summary(pca))
```

```{r, echo = FALSE}
plot(spca$importance[2, ] * 100, type = "b", bty = "n",
     ylab = "proportion of variance (%)", xlab = "PC")
```

### Pairs plot of PCs {.fragile}

```{box, type = 'info'}
PCs are not correlated.
```

```{r, echo = FALSE, fig.width = 8, fig.height = 4.5, out.width = '90%'}
dat_pca <- cbind(aq$Ozone, pca$x)
pairs(dat_pca, panel = panel, upper.panel = upper.panel,
      diag.panel = diag.panel, log = 1, cex = 0.7, pch = 20, gap = 0)
```

### Fit a linear model for PC1 {.fragile .allowframebreaks}

```{r}
fit3 <- lm(log10(aq$Ozone) ~ pca$x[, 1])
summary(fit3)
```

```{r, echo = FALSE, fig.width = 7, fig.height = 5, out.width = '90%'}
dat <- data.frame(Ozone = aq$Ozone, PC1 = pca$x[, 1])
ggplot(dat, aes(x = PC1, y = Ozone)) +
  geom_point() +
  stat_smooth(method = "lm", col = "red", formula = y ~ x) +
  scale_y_continuous(trans = log10_trans(),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)))
```

### Fit a LM for PC1 + PC3? {.fragile}

```{r, echo = FALSE}
fit4 <- lm(log10(aq$Ozone) ~ pca$x[, 1] + pca$x[, 3])
summary(fit4)
```

### Prediction {.fragile .allowframebreaks}

```{r, echo = FALSE, results = 'asis'}
p <- cbind(p, fit3 = 10^predict(fit2), fit4 = 10^predict(fit4))
dat <- reshape2::melt(p, id.vars = 1, variable.name = "Models",
                      value.name = "Ozone")
dfe <- cbind(dfe, fit3 = get_metrics(fit3), fit4 = get_metrics(fit4))
print.df(dfe)
```


```{r, echo = FALSE, fig.width = 7, fig.height = 3.3, out.width='90%'}
ggplot(dat, aes(n, Ozone, col = Models)) +
  geom_point(size = 0.5) + geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed")
```

```{r, echo = FALSE, fig.width = 7, fig.height = 3.5, out.width='90%'}
dat_res <- sapply(list(fit1, fit2, fit3, fit4), get_res)
colnames(dat_res) <- paste0("fit", seq_len(ncol(dat_res)))
dat_res <- as.data.frame(cbind(n = seq_len(nrow(dat_res)), dat_res))
dat_res <- reshape2::melt(dat_res, id.vars = 1, variable.name = "Models",
                          value.name = "Residuals")

ggplot(dat_res, aes(n, Residuals, col = Models)) +
  geom_point(shape = 1) +
  geom_hline(yintercept = 0, linetype = "dashed")

ggplot(dat_res, aes(Residuals, col = Models)) +
  stat_ecdf(geom = "step") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = 0.5, linetype = "dashed")
```

# References
