---
title: "Quantiles"
subtitle: "Fractiles, Percentiles"
abstract: "This is a presentation about Quantiles"
keywords: "quantile, percentile, R"
toc-multicol: false
---

```{r setup, include=FALSE}
source("setup.R")
n <- 11
```

# Definition
##
### Definition {.fragile}
<!-- .allowframebreaks -->

- A sample quantile, $q_p$, is a number having the same units as the data, which exceeds that proportion of the data given by the subscript $p$, with $0 < p < 1$.
- The sample quantile $q_p$ can be interpreted approximately as that value expected to exceed a randomly chosen member of the data set, with probability $p$.
- Equivalently, the sample quantile $q_p$ would be regarded as the $p \times 100th$ percentile of the data set [@Wilks2011].
- _Quantiles_ has the same unit with the data.

# Showcase
## How to calculate Quantiles
### How to calculate Quantiles {.fragile .allowframebreaks}

You measured height of $n = `r n`$ individuals.

```{r, echo=FALSE}
set.seed(1)
x <- round(runif(n, 50, 200), 1)
cat(paste0("x <- c(",  paste(x, sep = "", collapse = ","), ")"))
```

Sort the measurements:

```{r, echo=FALSE}
cat(paste0("x <- c(",  paste(sort(x), sep = "", collapse = ","), ")"))
```
\
What is $q_{0.5}$?
\
\
```{r, echo=FALSE, fig.width = 7, fig.height = 1.1, out.width = '90%'}
mark_quantile <- function(x, probs, col = "red") {
  qs <- quantile(x, probs)
  unlist(lapply(seq_len(length(qs)), function(i) {
    q <- qs[i]; p <- probs[i]
    label <- TeX(paste0("$q_{", p, "}$"), output = "character")
    list(geom_vline(xintercept = q, col = col, linetype = "dashed"),
      annotate("text", q, 0, label = label, col = col, vjust = -1.5,
               hjust = -0.1, parse = TRUE))
  }))
}
px <- pretty(x)
segs <- data.frame(x = px, xend = px, y = -0.02, yend = 0.02)
p1 <- ggplot(data.frame(x = x), aes(x)) + labs(x = "Length", y = "") +
  geom_hline(yintercept = 0, size = 0.3) + ylim(c(-0.1, 0.1)) +
  geom_segment(aes(x, y, xend = xend, yend = yend), data = segs,
               size = 0.3, inherit.aes = FALSE) +
  geom_point(y = 0, fill = "lightblue", shape = 21, size = 3) +
    theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) + mark_quantile(x, 0.5)
p1
```

\newpage
\
What are $q_{0.25}$ and $q_{0.75}$?

```{r, echo=FALSE, fig.width = 7, fig.height = 1.1, out.width = '90%'}
(p1 <- p1 + mark_quantile(x, c(0.25, 0.75), "darkgreen"))
```

```{box, type='info'}
$q_{0.25}$ and $q_{0.75}$ are also called as **Quartiles** (Hinges). Usually these are called the lower (LQ) and upper quartiles (UQ), respectively.
```

```{box, type='info'}
$q_{0.25}$ and $q_{0.75}$ can also be named as $25^{th}$ and $75^{th}$ **Percentiles.**
```

\newpage

What are $q_{0.1}$ and $q_{0.9}$?

```{r, echo=FALSE, fig.width = 7, fig.height = 1.1, out.width = '90%'}
(p1 <- p1 + mark_quantile(x, c(0.1, 0.9), "navyblue"))
```

What are $q_{0}$ and $q_{1}$?

```{r, echo=FALSE, fig.width = 7, fig.height = 1.1, out.width = '90%'}
(p1 <- p1 + mark_quantile(x, c(0, 1), "orange"))
```

# Interpretation
## 
### Interpretation {.fragile}

- $q_{0.5}$ seperates data into two equal parts, which is, $50\%$ of the data is below $q_{0.5}$ and $50\%$ of the data is above the $q_{0.5}$.
- $q_{0.25}$ seperates data into two parts, which is, $25\%$ of the data is below $q_{0.25}$ and $75\%$ of the data is above the $q_{0.25}$.
- $q_{0.9}$ seperates data into two parts, which is, $90\%$ of the data is below $q_{0.9}$ and $10\%$ of the data is above the $q_{0.9}$.

### quantile function in R {.fragile}

You can calculate quantiles by the `quantile` function.

```{r}
quantile(x, c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1))
```
\

```{box, type='alert'}
Be aware that $q_{0}$ and $q_{1}$ are equal to minimum and maximum of the data, respectively. 
```

# References
