---
title: "Summary Measures"
subtitle: "Location, Spread, Symmetry"
abstract: "This is a presentation about summary measures"
keywords: "location, spread, symmetry, mean, variance, standard deviation, R"
---

```{r setup, include=FALSE}
source("setup.R")
library(moments)
```

# Location
##
### Location {.fragile}
<!-- https://www.logicmatters.net/resources/pdfs/MathFonts.pdf -->
\begin{center}
\textbf{Mean:} (\emph{not robust to outliers})
\[
\bar{x} = \frac{1}{n} \sum_{i=1}^n{x_i}
\]

\textbf{Median:} (\emph{robust})
\[median=\begin{cases} x_{([n+1]/2)} &,n\;odd \\ \frac{x_{(n/2)}+x_{([n/2]+1)}}{2} &,n\; even \end{cases}\]

\textbf{Trimean:} (\emph{robust})
\[Trimean=\frac{q_{0.25} + 2q_{0.5} + q_{0.75}}{4}\]
\end{center}

See [@Wilks2011].

### Location in R {.fragile}

You have the following data. Let's calculate _mean_, _median_ and _trimean_.

```{r, echo=FALSE}
set.seed(1)
x1 <- sort(sample.int(50, 11))
x2 <- x1
x2[(length(x2) - 1):length(x2)] <- c(1000, 2000)
str <- paste0("x1 <- c(",  paste(x1, sep = "", collapse = ","), ")")
str <- paste(str, paste0("x2 <- c(",  paste(x2, sep = "", collapse = ","), ")"),
             sep = "\n")
cat(str)
```

_Trimean_ calculation can be done by the `trimean` function.

```{r}
trimean <- function(x) (2 * median(x) + sum(quantile(x, c(0.25, 0.75))))/4
cat(mean(x1), median(x1), trimean(x1))
cat(mean(x2), median(x2), trimean(x2))
```

# Spread
##
### Spread {.fragile}

\begin{center}
\textbf{Standard Deviation:} (\emph{not robust})
\[s=\sqrt{\frac{1}{n-1}\sum_{i=1}^n(x_i-\bar{x})^2}\]

\textbf{IQR:} (\emph{robust})
\[IQR =q_{0.75} - q_{0.25}\]

\textbf{Median Absolute Deviation:} (\emph{robust})
\[MAD=median\left| { x }_{ i }-{ q }_{ 0.5 } \right| \]
\end{center}

### Spread in R {.fragile}

Calculate _standard deviation_, _IQR_ and _MAD_ for $x_1$ and $x_2$.

```{r}
cat(sd(x1), IQR(x1), mad(x1))
cat(sd(x2), IQR(x2), mad(x2))
```

# Symmetry
##
### Symmetry {.fragile}

\begin{center}
\textbf{Skewness Coefficient:} (\emph{not robust})
\[\gamma =\frac{ \frac{1}{n} \sum_{i=1}^n{{(x_i - \bar{x})}^3}}{s^3}\]

\textbf{Yule-Kendall Index:} (\emph{robust})
\[\gamma_{yk} = \frac{q_{0.25} - 2q_{0.5} + q_{0.75}}{IQR} \]
\end{center}

### Symmetry in R {.fragile}

Calculate _skewness coefficient_ and _YK-index_  for $x_1$ and $x_2$.

```{r}
library(moments)
yk <- function(x) (sum(quantile(x, c(0.25, 0.75))) - 2 * median(x))/IQR(x)
cat(skewness(x1), yk(x1))
cat(skewness(x2), yk(x2))
```

# References
