---
title: "t-test"
subtitle: "One Sample t-test"
abstract: "This is a presentation about t-test"
keywords: "hypothesis testing, t-test, R"
---

```{r setup, include=FALSE}
knitr_cache <- TRUE
source("setup.R")
n <- 20
df <- n - 1
```

# Basics

## Set up \hypt{}
### Five Steps

1. _\ts{}_ is mean. Compare sample mean ($\bar{x}$) against population mean ($\mu$).
2. $H_0$: $\bar{x}$ is equal to $\mu$. ($\bar{x} = \mu$), ($\pval > \alpha$).
3. $H_A$: $\bar{x}$ is different than $\mu$. ($\bar{x} \neq \mu$), ($\pval \le \alpha$)
4. _null distribution_ is _student-t distribution_, which is simply the sampling distribution for the test statistic, if the $H_{0}$ is _true_.
5. Compare the observed _\ts{}_ to the _null distribution_ [@Wilks2011].


## $\alpha$ and Confidence Levels
### $\alpha$ and Confidence Levels {.fragile}

1. **Significance ($\boldsymbol\alpha$) level:** The probability of rejecting the $H_0$ when the null hypothesis is true.
2. **Confidence Level:** The probability of confidence interval contains the true population mean ($\mu$). \textcolor{violet}{It is not the probability of which the $\mu$ lies within interval.}

```{box, type='equation', latex.options='[width=\\linewidth/3]'}
$\alpha = 0.05$

$conf.level = 1- \alpha$
```

## When to use?
### When to use?

You want to find out;

1. if your observations are from a _known normal distribution_ (which $\mu$ is known).
2. if your observations are similar to the observations (which $\bar{y}$ is known) from a previous study.

In both cases, you need to know $\mu$ or $mean(\bar{y})$ of previous observations.

# Define Question
### Define Question {.fragile}

Assume that you know the mean age of a female population;

```{r}
mu <- 33.2 # Mean age of female population
```

You _randomly_ asked $`r n`$ females for their age and you get the answers below;

```{r, echo=FALSE, style="nonum"}
set.seed(3)
x <- round(rnorm(n, 34, 15))
mx <- mean(x)
cat(paste0("x <- c(",  paste(x, sep = "", collapse = ","), ")"))
```

```{box, type='question'}
Are they from a normal distribution which $\mu = 33.2$?
```

### Kernel Density plot

```{r fig1_kernel_density, echo=FALSE,out.width = '60%'}
labels <- c(sprintf("$\\bar{x} = %.1f$", mx),
            sprintf("$s = %.1f$", sd(x)),
            sprintf("$df = %d - 1 = %d$", n, df),
            "$\\mu$")
p1 <- ggplot(data.frame(x), aes(x = x)) +
  labs(x = "Age of females", y = "Density") +
  geom_density() + xlim(c(0, 100)) +
  geom_vline(xintercept = c(mean(x), mu), linetype = "dashed",
             col = c("black", "red")) +
  annotate("text", x = c(60, 60, 58.7, mu + 1),
           y = c(0.024, 0.0215, 0.019, 0.005), hjust = 0,
           label = TeX(labels, output = "character"), parse = TRUE,
           col = c("black", "black", "black", "red")) +
  geom_rug()
print(p1)
```

```{box, type='question'}
Don't we have any female older than 75?
```

# How to Solve?
## Define $\alevel$
### Define $\alevel$ {.fragile}

First, you need to define $\alevel$;

```{r}
alpha <- 0.05
conf_level <- 1 - alpha
```

## Calculate SE and t-value
### Calculate SE and t-value {.fragile}

Calculate _Standard Error_ and $\tval$;

```{r}
n <- length(x)
se <- sd(x) / sqrt(n)
t <- (mean(x) - mu) / se # What's happening here?
```
```{r, echo = FALSE}
cat(sprintf("SE = %.2f\tt = %.2f", se, t))
```

```{box, type='question'}
1- Why $SE$ and $\sigma$ are different?

2- What is the meaning of $t$?
```
### student-t distribution {.fragile}

Null distribution is _student-t distribution_ with degrees of freedom $df = `r df`$.

```{r fig2_tdist, echo=FALSE}
y <- seq(-5, 5, 0.01)
dat_st <- data.frame(x = y, y = dt(y, df))
rx <- range(y)
p2 <- ggplot(dat_st, aes(x, y)) +
  labs(x = "Quantiles", y = "Density") +
  geom_line(col = "orange", size = 1) +
  geom_vline(xintercept = c(0, t), linetype = "dashed",
             col = c("black", "violet")) +
  annotate("text", x = 0.3, y = 0.05,
           label = TeX("$\\mu'$", output = "character"),
           parse = TRUE) +
    annotate("text", x = 1.7, y = 0.2, label = "null distribution",
           col = "orange", angle = -76) +
  annotate("text", x = t, y = 0.35, label = "t", hjust = 2,
           col = "violet")
print(p2)
```

```{box, type='alert'}
student-t distribution is defined only by degrees of freedom ($\mu = 0$ and $\sigma = 1$).
```

# Solution by Quantiles of $\alevel$
## Calculate Quantiles of $\alevel$
### Calculate Quantiles of $\alevel$ {.fragile}

We will use _student-t distribution_ to calculate quantiles of $\alevel$ at both tails.

```{r}
q <- qt(c(alpha / 2, 1 - alpha / 2), df)
```

Mark the quantiles of $\alevel$ on the plot.

\begin{center}
$q_{\alpha} = `r round(abs(q[1]), 3)`$
\end{center}

```{r fig3_t_quantiles, echo = FALSE, out.width = '40%'}
qlabels <- TeX(c("$-q_{\\alpha}$", "$q_{\\alpha}$"),
               output = "character")
qdata <- data.frame(x = q, y = 0, xend = q, yend = dt(q, df))
p3 <- p2 +
  geom_segment(aes(x, y, xend = xend, yend = yend), qdata,
               col = "red", linetype = 3, inherit.aes = FALSE) +
  geom_text(aes(x = x, y = y), qdata, label = qlabels,
                  col = "red", parse = TRUE, vjust = 1) +
    geom_ribbon(data = subset(dat.st, x >= rx[1] & x <= q[1]),
              aes(ymax = y), ymin = 0, fill = "yellow", colour = NA,
              alpha = 0.5) +
  geom_ribbon(data = subset(dat.st, x >= q[2] & x <= rx[2]),
              aes(ymax = y), ymin = 0, fill = "yellow", colour = NA,
              alpha = 0.5)
print(p3)
```

## Decide test result
### Method-I (Decide test result)

You can use $t$ value and $q_{\alpha}$ values to decide the test result. 

- If $-q_{\alpha} > t < q_{\alpha}$, **fail to reject** $H_0$.
- If $-q_{\alpha} \le t \ge q_{\alpha}$, reject $H_0$.

```{r, echo = FALSE, out.width = '50%'}
include_graphics("_figures/t_test-fig3_t_quantiles-1.pdf")
```

# Solution by Confidence Interval
## Calculate Confidence Interval
### Calculate Confidence Interval {.fragile}

We know the quantiles of $\alevel$ for _student-t distribution_, so we can transform the quantiles back to original values for $\bar{x}$. Hence;

```{box, type='equation', latex.options='[width=0.4\\linewidth]'}
$CI = \bar{x} + (-q_{\alpha}, q_{\alpha}) \times SE$
```

```{r fig4_conf_int, echo = FALSE}
(ci <- mean(x) + q * se)
p4 <- p1 + geom_rect(xmin = ci[1], xmax = ci[2], ymin = -Inf,
                     ymax = Inf, alpha = 0.02, fill = "forestgreen")
print(p4)
```

### Deep look into Confidence Interval

```{box, type='question'}
How did we calculate the confidence interval? What did we use?
```

```{box, type='equation', latex.options='[width=0.7\\linewidth]'}
$CI(\bar{x}, q_{\alpha}, SE)$, 
$SE(s, n)$
$q_{\alpha}(\alpha, df)$,
$df(n)$
```

```{box, type='equation', latex.options='[width=0.2\\linewidth]'}
$CI(\bar{x}, s, n, \alpha)$
```

1. If $\alpha \rightarrow 1$, $q_{\alpha} \rightarrow 0$, $CI \rightarrow \bar{x}$ (We want to keep $\alpha$ low).
2. If $n \rightarrow \infty$, $SE \rightarrow 0$, $CI \rightarrow \bar{x}$ (Law of Large Numbers).
3. If $s \rightarrow \infty$, $SE \rightarrow \infty$, $CI \rightarrow \infty$.
4. If $\bar{x} \rightarrow \pm \infty$, $CI \rightarrow \pm \infty$ (Shifts the location).

## Decide test result
### Method-II (Decide test result) {.fragile}

You can use $\mu$ and confidence interval values to decide the test result. 

- If $c_1 > \mu < c_2$, **fail to reject** $H_0$.
- If $c_1 \le \mu \ge c_2$, reject $H_0$.

```{r, echo = FALSE, out.width = '50%'}
include_graphics("_figures/t_test-fig4_conf_int-1.pdf")
```


# Solution by $\pval$
## $\pval$
### What is $\pval$ {.fragile}

$\pval$ is the \textcolor{violet}{probability which generates the data}, \textcolor{blue}{or something else which is equal} \textcolor{red}{or rarer}.

\begin{center}
$t = `r round(t, 2)`$
\end{center}

```{r fig5_pval1, echo = FALSE}
p_val <- pt(-abs(t), df)
p5 <- p2 + geom_vline(xintercept = -t, linetype = "dashed",
                      col = "blue") +
  annotate("text", x = -t, y = 0.35, label = c("-t"),
           hjust = -0.5, col = c("blue")) +
  geom_ribbon(data = subset(dat.st, x >= rx[1] & x <= -abs(t)),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA,
              alpha = 0.5) +
  geom_ribbon(data = subset(dat.st, x >= abs(t) & x <= rx[2]),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA,
              alpha = 0.5)
print(p5)
```

### How to calculate $\pval$ in R? {.fragile}

```{r}
p_value <- 2 * pt(-abs(t), df)
```
\begin{center}
$\pval = p_1 + p_2 = `r sprintf("%.3f + %.3f = %.3f", p_val, p_val, p_value)`$
\end{center}

```{r fig6_pval2, echo = FALSE}
p6 <- p5 +
  annotate("segment", x = c(-3.5, 3.5), y = 0.1,
             xend = c(-1.5, 1.5), yend = 0.05,
             arrow = arrow(angle = 20, length = unit(2, "mm"),
                           type = "closed"), col = "brown") +
    annotate("label", x = c(-3.5, 3.5), y = 0.1,
           label = TeX(sprintf("$p_%d = %.3f$", 1:2, p_val),
                       output = "character"), parse = TRUE, col = "brown")
print(p6)
```

```{box, type='question'}
How to calculate $\pval$ for normal distribution?
```

## Decide test result
### Method-III (Decide test result) {.fragile}

You can use $\pval$ to decide the test result. 

- If $\pval > \alpha$, **fail to reject** $H_0$.
- If $\pval \le \alpha$, reject $H_0$.

```{box, type = 'info'}
As we **failed to reject** $H_0$ by three methods, we can conclude that our observations ($\bar{x}=`r round(mean(x), 2)`$) come from a normal distribution which $\mu = `r mu`$ with $`r 100 * conf.level`\%$ confidence. The means are not significantly different. Also, confidence interval is $CI \cong [`r sprintf("%.2f", ci)`]$.
```

# t-test function in R
## Use t-test for \hypt

The steps above can be easly done by `t.test` funcion in R.

```{box, type='equation', latex.options='[width=0.7\\linewidth]'}
$`r sprintf("t = %.3f, df = %d, \\pval = %.3f", t, df, p_value)`$

$CI = (`r sprintf("%.2f", ci)`)$
```

### _Equality_ for mean (two-tail) {.fragile}

1. $H_{0}$: $\bar{x}$ is equal to $\mu$. ($\bar{x} = \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is not equal to $\mu$. ($\bar{x} \ne \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, mu = mu, conf.level = conf.level)
```

### _Less than_ for mean (right-tail) {.fragile}

1. $H_{0}$: $\bar{x}$ is equal or greater than $\mu$. ($\bar{x} \ge \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is less than $\mu$. ($\bar{x} < \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "less", mu = mu, conf.level = conf.level)
```

### _Greater than_ for mean (left-tail) {.fragile}

1. $H_{0}$: $\bar{x}$ is equal or less than $\mu$. ($\bar{x} \le \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is greater than $\mu$. ($\bar{x} > \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "greater", mu = mu, conf.level = conf.level)
```

## One-Tail or Two Tail?
### Which one should I use?{.fragile}

```{box, type='question'}
Which does it make sense? 

* Equality
* Less
* Greater
```

# What if \hn{} is false?
## Another t-test Example
### _Equality_ for mean (two-tail) {.fragile}

What if $\mu = 37$? Be aware confidence interval.

1. $H_{0}$: $\bar{x}$ is equal to $\mu$. ($\bar{x} = \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is not equal to $\mu$. ($\bar{x} \ne \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, mu = 37, conf.level = conf.level)
```

### _Less than_ for mean (right-tail) {.fragile}

1. $H_{0}$: $\bar{x}$ is equal or greater than $\mu$. ($\bar{x} \ge \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is less than $\mu$. ($\bar{x} < \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "less", mu = 37, conf.level = conf.level)
```

### _Greater than_ for mean (left-tail) {.fragile}

1. $H_{0}$: $\bar{x}$ is equal or less than $\mu$. ($\bar{x} \le \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is greater than $\mu$. ($\bar{x} > \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "greater", mu = 37, conf.level = conf.level)
```

# Interpret p-value
### How to interpret p-value? {.fragile}

Even if $H_0$ is \textcolor{red}{$true$}, there is still $5\%$ chance to reject $H_0$.
```{r}
try <- 1000
pvals <- sapply(1:try, function(i) {
  set.seed(i)
  t.test(rnorm(n))$p.value
})
per_of_type1 <- 100 * sum(pvals < 0.05) / try
```

```{r fig7_pval_hist, echo = FALSE, fig.height = 1.5}
cat(sprintf("Percent of Type I Error : %1.f%%", per_of_type1))
breaks <- seq(0, 1, 0.05)
a <- c(1, rep(NA, length(breaks) - 2))
per <- function(x) {
  c(percent(x[1] / length(pvals)), rep("", length(x) - 1))
}
ggplot(data.frame(x = pvals), aes(x)) +
  geom_histogram(breaks = breaks, col = "black",
                 fill = c("red", rep(NA, length(breaks) - 2)), alpha = 0.5) +
  stat_bin(breaks = breaks, aes(label = per(..count..)),
           vjust = -0.5, geom = "text", na.rm = TRUE) +
  labs(x = "p-values")

# power.t.test(n, mu - mean(x), sd(x), alpha, type = "one.sample")
```

# References

