---
title: "Hypothesis Testing"
subtitle: "(t-test)"
abstract: "This is a presentation about Hypothesis Testing"
keywords: "hypothesis testing, t-test, R"
titlegraphic: images/itu2.eps
titlegraphic-width: 0.15
author:
  - \href{mailto:sezenismail@gmail.com}{İ.SEZEN}
  - \href{mailto:alper.unal@gmail.com}{A.ÜNAL}
institute:
  - Eurasia Institute of Earth Sciences
institute-short: İTÜ
date: \today
fontsize: 10pt
navigation: horizontal
header-includes:
  - \usepackage{ragged2e}
  - \apptocmd{\frame}{}{\justifying}{}
section-titles: # Hide section title pages
colorlinks: yes
toc: yes
toc-title: Summary
toc-depth: 3
justifying: yes
output:
  beamer_presentation:
    template: default.beamer
    keep_tex: yes
    slide_level: 3
    fig_caption: yes
    theme: CambridgeUS
    colortheme: euresia
    highlight: tango
    pandoc_args: ["--listings", "-H", "listings-setup.tex"]
    # pandoc_args: ["--highlight-style", "my.theme"]
    includes:
      in_header: before_body.tex
    #   after_body: after_body.tex
    
---

```{r setup, include=FALSE}
library(kableExtra)
options(knitr.table.format = "latex")
knitr::opts_chunk$set(echo = FALSE, background = '#FEF8DF')
# a common hook for messages, warnings and errors
optlist <- c("caption", "numbers", "basicstyle", "firstnumber", "style")
hook <- function(x, options, ..., begin = "lstlisting") {
  if (length(x) > 1) x <- paste(x, collapse = '\n')
  opts <- options[optlist]
  opts <- append(list(...), opts)
  opts[sapply(opts, is.null)] <- NULL
  opts <- ifelse(length(opts) < 1, "",
                 sprintf("[%s]",paste0(names(opts), "=", opts, collapse = ",")))
  knitr::raw_latex(
    sprintf("\n\\begin{%s}%s\n%s\n\\end{%s}\n",begin, opts, x, begin))
}
hook_lst_bf = function(x, options) {
  hook(x, options, basicstyle = "{\\bfseries}")
}
knitr::knit_hooks$set(source = function(x, options) {
  if (length(x) > 1) x <- paste(x, collapse = '\n')
  hook(x, options, begin = "rcode")
}, output = function(x, options) {
  if ("caption" %in% names(options)) options$caption = "\\mbox{}"
  options$firstnumber = 1
  x <- substr(x, 1, nchar(x) - 2)
  if (substr(x, 1, 4) == paste(options$comment,"\n")) x <- substr(x, 5, nchar(x))
  hook(x, options, begin = "outp")
}, warning = hook_lst_bf, message = hook_lst_bf, error = hook_lst_bf)

## empty highlight header since it is not useful any more
knitr::set_header(highlight = "")
```

# \hypt
## What is \hypt?
### What is \hypt?

\hyp\ is,

- a supposition or proposed explanation made on the basis of limited evidence 
  as a starting point for further investigation.
- a testable statement about the relationship **between two or more variables**.
- a proposed explanation for some observed phenomenon^[[1]]^.
- an assumption about a population parameter. This assumption may or may not be 
  true (For instance, a population mean is equal to $10$)^[[2]]^.

[1]: https://whatis.techtarget.com/definition/hypothesis
[2]: https://stattrek.com/statistics/dictionary.aspx?definition=statistical_hypothesis

### What is \hypt?

\hyp\ is,

- a statistical method that is used in making statistical decisions using 
  experimental data^[[3]]^.
- basically an assumption that we make about the population parameter^[[3]]^.
- is a method of [statistical inference](https://en.wikipedia.org/wiki/Statistical_inference)^[[4]]^.
- The theory, methods, and practice of testing a \hyp\ by comparing it 
  with the **\hn{}**.

[3]: https://www.statisticssolutions.com/hypothesis-testing/
[4]: https://en.wikipedia.org/wiki/Statistical_hypothesis_testing

### Usage Cases

- A _sample_ is compared against a population parameter.
- Two statistical data sets are compared^[[3]]^.
- a data set obtained by sampling is compared against a synthetic data set 
  from an idealized model^[[3]]^.
  
### Steps of \hyp\ test

1. Identify a _\ts{}_ that is appropriate to the data and question.
2. Define a **\hn{}**, $H_{0}$.
3. Define an **\ha{}**, $H_{A}$.
4. Obtain the _null distribution_, which is simply the sampling distribution
  for the test statistic, if the _\hn{}_ is _true_.
5. Compare the observed _\ts{}_ to the _null distribution_.

### t-test

1. $H_{0}$: Mean is equal to $\mu$ $\mu - \bar{x} = 0$ (${p-value} < \alpha$)
2. $H_{A}$: Mean is different than $\mu$ $\mu - \bar{x} \neq 0$ ($p-value \ge \alpha$)

### Confidence and Alpha Levels {.fragile .allowframebreaks}

1. **Significance ($\alpha$) level:** The probability of rejecting the $H_{0}$ when the null hypothesis is true.
2. **Confidence Level:** The probability of confidence interval contains the true population mean ($\mu$). \textcolor{violet}{It is not the probability of which the $\mu$ lies within interval.}

```{r mycode1, echo = TRUE, caption="T-test", firstnumber=1}
alpha <- 0.05
conf.level <- 1 - alpha
```
