---
title: "Hypothesis Testing"
subtitle: "(t-test)"
abstract: "This is a presentation about Hypothesis Testing"
keywords: "hypothesis testing, t-test, R"
titlegraphic: images/itu2.eps
titlegraphic-width: 0.15
author:
  - \href{mailto:sezenismail@gmail.com}{İ.SEZEN}
  - \href{mailto:alper.unal@gmail.com}{A.ÜNAL}
institute:
  - Eurasia Institute of Earth Sciences
institute-short: İTÜ
date: \today
fontsize: 10pt
navigation: horizontal
# header-includes:
#   include-before:
# include-after: \input{after_body.tex}
section-titles: # Hide section title pages
colorlinks: yes
toc: yes
toc-title: Summary
toc-depth: 3
justifying: yes
output:
  beamer_presentation:
    template: default.beamer
    keep_tex: yes
    slide_level: 3
    fig_caption: yes
    theme: CambridgeUS
    colortheme: euresia
    highlight:
    pandoc_args: ["--listings", "-H", "listings-setup.tex"]
    # pandoc_args: ["--highlight-style", "my.theme"]
    includes:
      in_header: before_body.tex
    #   after_body: after_body.tex
    
---

```{r setup, include=FALSE}
library(kableExtra)
options(knitr.table.format = "latex")
knitr::opts_chunk$set(echo = FALSE, background = '#FEF8DF')
# a common hook for messages, warnings and errors
optlist <- c("caption", "numbers", "basicstyle", "firstnumber", "style")
hook <- function(x, options, ..., begin = "lstlisting") {
  if (length(x) > 1) x <- paste(x, collapse = '\n')
  opts <- options[optlist]
  opts <- append(list(...), opts)
  opts[sapply(opts, is.null)] <- NULL
  opts <- ifelse(length(opts) < 1, "",
                 sprintf("[%s]",paste0(names(opts), "=", opts, collapse = ",")))
  knitr::raw_latex(
    sprintf("\n\\begin{%s}%s\n%s\n\\end{%s}\n",begin, opts, x, begin))
}
hook_lst_bf = function(x, options) {
  hook(x, options, basicstyle = "{\\bfseries}")
}
knitr::knit_hooks$set(source = function(x, options) {
  if (length(x) > 1) x <- paste(x, collapse = '\n')
  hook(x, options, begin = "rcode")
}, output = function(x, options) {
  if ("caption" %in% names(options)) options$caption = "\\mbox{}"
  options$firstnumber = 1
  x <- substr(x, 1, nchar(x) - 2)
  if (substr(x, 1, 4) == paste(options$comment,"\n")) x <- substr(x, 5, nchar(x))
  hook(x, options, begin = "outp")
}, warning = hook_lst_bf, message = hook_lst_bf, error = hook_lst_bf)

## empty highlight header since it is not useful any more
knitr::set_header(highlight = "")
```

# \hypt
## What is \hypt?
### What is \hypt?

\hyp\ is,

- a supposition or proposed explanation made on the basis of limited evidence 
  as a starting point for further investigation.
- a testable statement about the relationship **between two or more variables**.
- a proposed explanation for some observed phenomenon^[[1]]^.
- an assumption about a population parameter. This assumption may or may not be 
  true (For instance, a population mean is equal to $10$)^[[2]]^.

[1]: https://whatis.techtarget.com/definition/hypothesis
[2]: https://stattrek.com/statistics/dictionary.aspx?definition=statistical_hypothesis

### What is \hypt?

\hyp\ is,

- a statistical method that is used in making statistical decisions using 
  experimental data^[[3]]^.
- basically an assumption that we make about the population parameter^[[3]]^.
- is a method of [statistical inference](https://en.wikipedia.org/wiki/Statistical_inference)^[[4]]^.
- The theory, methods, and practice of testing a \hyp\ by comparing it 
  with the **\hn{}**.

[3]: https://www.statisticssolutions.com/hypothesis-testing/
[4]: https://en.wikipedia.org/wiki/Statistical_hypothesis_testing

### Usage Cases

- A _sample_ is compared against a population parameter.
- Two statistical data sets are compared^[[3]]^.
- a data set obtained by sampling is compared against a synthetic data set 
  from an idealized model^[[3]]^.
  
### Steps of \hyp\ test

1. Identify a _\ts{}_ that is appropriate to the data and question.
2. Define a **\hn{}**, $H_{0}$.
3. Define an **\ha{}**, $H_{A}$.
4. Obtain the _null distribution_, which is simply the sampling distribution
  for the test statistic, if the _\hn{}_ is _true_.
5. Compare the observed _\ts{}_ to the _null distribution_.

## Example #1 {.fragile}

```{r code1, echo=TRUE, caption = "Hypothesis Testing"}
# utility functions utility functions utility functions
readinteger <- function(...) {
  a <- data.frame(matrix(1:20, ncol = 12, nrow = 20))
  ncol(a[,12:23])
  z <- median(a)
  x <- 2^35; integer(c(4321,3:56)) -> y; c <- NA
  a <<- x == y & x != y & x <= y | x >= y || x < y && x > y
  b <- !x + y - 1.5 * 12 / 0.3 %% 2 %/% 12 %*% 5 %in% a %o% x
  s <- "^[0-9]+$?"
  n <- !rnorm(x)
  plot(x, y, type = "l")
  if (length(n) == x) print(s)
  warning("message warning")
  stopifnot(1 == 1, all.equal(pi, 3.14159265), 1 < 2, 
            LETTERS, month.abb, airquality) # all TRUE
  return(as.integer(n))
}
```


### Slide with Plot

```{r pressure1}
plot(pressure)
```

## Sub-Section 2
### Frame Title 1
justify This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

# Section 2
## Sub-Section 1
### Frame Title 1
justify This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

