---
title: "t-test"
subtitle: ""
abstract: "This is a presentation about t-test"
keywords: "hypothesis testing, t-test, R"
titlegraphic: images/itu2.eps
titlegraphic-width: 0.15
author:
  - \href{mailto:sezenismail@gmail.com}{İ.SEZEN}
  - \href{mailto:alper.unal@gmail.com}{A.ÜNAL}
institute:
  - Eurasia Institute of Earth Sciences
institute-short: İTÜ
date: \today
fontsize: 10pt
navigation: horizontal
header-includes:
  - \usepackage{ragged2e}
  - \apptocmd{\frame}{}{\justifying}{}
beamerconfig:
    blocks: # default rounded, no shadow. Do not recommend shadows.
      shadow:
section-titles: # Hide section title pages
colorlinks: yes
toc: yes
toc-title: Summary
toc-depth: 3
justifying: yes
output:
  beamer_presentation:
    template: default.beamer
    keep_tex: yes
    slide_level: 3
    fig_caption: yes
    theme: CambridgeUS
    colortheme: euresia
    highlight: tango
    pandoc_args: ["--listings", "-H", "listings-setup.tex"]
    # pandoc_args: ["--highlight-style", "my.theme"]
    includes:
      in_header: before_body.tex
      # after_body: after_body.tex
    
---

```{r setup, include=FALSE}
library(scales)
library(latex2exp)
library(ggplot2)
library(ggrepel)
library(kableExtra)
global_theme <- ggthemes::theme_pander() +
  theme(rect = element_rect(fill = "transparent"),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA))
theme_set(global_theme)
options(knitr.table.format = "latex")
knitr::opts_chunk$set(echo = TRUE, background = '#FEF8DF', 
                      fig.align = "center", fig.width = 4,
                      fig.height = 3, out.width = '50%')
# a common hook for messages, warnings and errors
optlist <- c("caption", "numbers", "basicstyle", "firstnumber", "style")
hook <- function(x, options, ..., begin = "lstlisting") {
  if (length(x) > 1) x <- paste(x, collapse = '\n')
  opts <- options[optlist]
  opts <- append(list(...), opts)
  opts[sapply(opts, is.null)] <- NULL
  opts <- ifelse(length(opts) < 1, "",
                 sprintf("[%s]",paste0(names(opts), "=", opts, collapse = ",")))
  knitr::raw_latex(
    sprintf("\n\\begin{%s}%s\n%s\n\\end{%s}\n",begin, opts, x, begin))
}
hook_lst_bf = function(x, options) {
  hook(x, options, basicstyle = "{\\bfseries}")
}
knitr::knit_hooks$set(source = function(x, options) {
  if (length(x) > 1) x <- paste(x, collapse = '\n')
  hook(x, options, begin = "rcode")
}, output = function(x, options) {
  if ("caption" %in% names(options)) options$caption = "\\mbox{}"
  options$firstnumber = 1
  x <- substr(x, 1, nchar(x) - 1)
  if (substr(x, 1, 4) == paste(options$comment,"\n")) x <- substr(x, 5, nchar(x))
  hook(x, options, begin = "outp")
}, warning = hook_lst_bf, message = hook_lst_bf, error = hook_lst_bf)

## empty highlight header since it is not useful any more
knitr::set_header(highlight = "")
```

# t-test

## One Sample t-test
### Steps of t-test

1. _\ts{}_ is mean. Compare sample mean ($\bar{x}$) against population mean ($\mu$).
2. $H_0$: $\bar{x}$ is equal to $\mu$. ($\bar{x} = \mu$), ($\pval > \alpha$).
3. $H_A$: $\bar{x}$ is different than $\mu$. ($\bar{x} \neq \mu$), ($\pval \le \alpha$)
4. _null distribution_ is _student-t distribution_, which is simply the sampling distribution for the test statistic, if the $H_{0}$ is _true_.
5. Compare the observed _\ts{}_ to the _null distribution_.

### Confidence and Alpha Levels {.fragile}

1. **Significance ($\boldsymbol\alpha$) level:** The probability of rejecting the $H_0$ when the null hypothesis is true.
2. **Confidence Level:** The probability of confidence interval contains the true population mean ($\mu$). \textcolor{violet}{It is not the probability of which the $\mu$ lies within interval.}

```{r mycode1, caption="t-test", firstnumber=1}
alpha <- 0.05
conf.level <- 1 - alpha
```

### When to use? {.fragile .allowframebreaks}

You want to find out;

1. if your observations come from a known normal distribution (which $\mu$ is known).
2. if your observations are similar to the observations (which $\bar{y}$ is known) from a previous study.

In both cases, you need to know $\mu$ or $mean(\bar{y})$ of previous observations.

\begin{center}\rule{0.5\linewidth}{0.5000000000pt}\end{center}

Assume that you know the mean age of female population;

```{r}
mu <- 33.2 # Mean age of female population
```

\newpage

```{r, echo=FALSE}
n <- 20
```

You _randomly_ asked $`r n`$ females their age and you get the results below;

```{r mycode2, echo=FALSE, style="nonum"}
df <- n - 1
set.seed(3)
x <- round(rnorm(n, 34, 15))
mx <- mean(x)
labels <- c(sprintf("$\\bar{x} = %.1f$", mx),
            sprintf("$s = %.1f$", sd(x)),
            sprintf("$df = %d - 1 = %d$", n, df),
            "$\\mu$") 
cat(paste(x, sep = "", collapse = ","))
p1 <- ggplot(data.frame(x), aes(x = x)) + 
  labs(x = "Age of females", y = "Density") +
  geom_density() + xlim(c(0, 100)) +
  geom_vline(xintercept = c(mean(x), mu), linetype = "dashed",
             col = c("black", "red")) +
  annotate("text", x = c(60, 60, 58.7, mu + 1), 
           y = c(0.024, 0.0215, 0.019, 0.005), hjust = 0,
           label = TeX(labels, output = "character"), parse = TRUE,
           col = c("black", "black", "black", "red")) +
  geom_rug()
print(p1)
```

### Standard Error and t-value {.fragile .allowframebreaks}

Are they coming from a normal distribution which $\mu = 33.2$?

\begin{center}\rule{0.5\linewidth}{0.5000000000pt}\end{center}

Let's calculate _Standard Error_ and _t-value_;

```{r}
n <- length(x)
se <- sd(x)/sqrt(n)
t <- (mean(x) - mu)/se # What's happening here?
```
```{r, echo = FALSE}
cat(sprintf("SE = %.2f\tt = %.2f", se, t))
```

\begin{exampleblock}{Question}
- Why $SE$ and $\sigma$ are different?

- What is the meaning of $t$?
\end{exampleblock}

\newpage

Null distribution here is student-t distribution with degrees of freedom $df = `r n - 1`$.

```{r, echo=FALSE}
y <- seq(-5, 5, 0.01)
dat.st <- data.frame(x = y, y = dt(y, df))
rx <- range(y)
p2 <- ggplot(dat.st, aes(x, y)) + 
  labs(x = "Quantiles", y = "Density") +
  geom_line(col = "orange", size = 1) +
  geom_vline(xintercept = c(0, t), linetype = "dashed", 
             col = c("black", "violet")) +
  annotate("text", x = 0.3, y = 0.05, 
           label = TeX("$\\mu'$", output = "character"), 
           parse = TRUE) +
    annotate("text", x = 1.7, y = 0.2, label = "null distribution", 
           col = "orange", angle = -76) +
  annotate("text", x = t, y = 0.35, label = "t", hjust = 2,
           col = "violet")
print(p2)
```

\begin{alertblock}{Alert}
student-t distribution is defined only by degrees of freedom ($\mu = 0$ and $\sigma = 1$).
\end{alertblock}

### Calculate Quantiles of $\alpha$ Level {.fragile}

We will use _student-t distribution_ to calculate quantiles of $\alpha$ level at both tails.

```{r}
q <- qt(c(alpha/2, 1 - alpha/2), df)
```

Let's mark the quantiles of $\alpha$ level on the plot.

\begin{center}
$q_{\alpha} = `r round(abs(q[1]), 3)`$
\end{center}

```{r, echo = FALSE, out.width = '40%'}
qlabels <- TeX(c("$-q_{\\alpha}$", "$q_{\\alpha}$"), 
               output = "character")
qdata <- data.frame(x = q, y = 0, xend = q, yend = dt(q, df))
p3 <- p2 + 
  geom_segment(aes(x, y, xend = xend, yend = yend), qdata,
               col = "red", linetype = 3, inherit.aes = FALSE) +
  geom_text(aes(x = x, y = y), qdata, label = qlabels,
                  col = "red", parse = TRUE, vjust = 1) +
    geom_ribbon(data = subset(dat.st, x >= rx[1] & x <= q[1]),
              aes(ymax = y), ymin = 0, fill = "yellow", colour = NA,
              alpha = 0.5) +
  geom_ribbon(data = subset(dat.st, x >= q[2] & x <= rx[2]),
              aes(ymax = y), ymin = 0, fill = "yellow", colour = NA,
              alpha = 0.5)
print(p3)
```

### Method-I (Decide tests result) {.fragile}

You can use $t$ value and $q_{\alpha}$ values to decide the test result. 

- If $-q_{\alpha} > t < q_{\alpha}$, **fail to reject** $H_0$.
- If $-q_{\alpha} \le t \ge q_{\alpha}$, reject $H_0$.

```{r, echo = FALSE}
print(p3)
```

### Calculate Confidence Interval {.fragile .allowframebreaks}

We know the quantiles of $\alpha$ level for _student-t distribution_, so we can transform the quantiles back to original values for $\bar{x}$. Hence;

\begin{center}
$CI = \bar{x} + (-q_{\alpha}, q_{\alpha}) \times SE$
\end{center}

```{r, echo = FALSE}
(ci <- mean(x) + q * se)
p4 <- p1 + geom_rect(xmin = ci[1], xmax = ci[2], ymin = -Inf, 
                     ymax = Inf, alpha = 0.02, fill = "red")
print(p4)
```
\newpage

\begin{alertblock}{Question}
How did we calculate the confidence interval? What did we use?
\end{alertblock}

\begin{center}
$CI(\bar{x}, q_{\alpha}, SE)$, 
$SE(s, n)$
$q_{\alpha}(\alpha, df)$,
$df(n)$

$CI(\bar{x}, s, n, \alpha)$
\end{center}

1. If $\alpha \rightarrow 1$, $q_{\alpha} \rightarrow 0$, $CI \rightarrow \bar{x}$ (We want to keep $\alpha$ low).
2. If $n \rightarrow \infty$, $SE \rightarrow 0$, $CI \rightarrow \bar{x}$ (Law of Large Numbers).
3. If $s \rightarrow \infty$, $SE \rightarrow \infty$, $CI \rightarrow \infty$.
4. If $\bar{x} \rightarrow \pm \infty$, $CI \rightarrow \pm \infty$ (Shifts the location).

### Method-II (Decide tests result) {.fragile}

You can use $\mu$ and confidence interval values to decide the test result. 

- If $c_1 > \mu < c_2$, **fail to reject** $H_0$.
- If $c_1 \le \mu \ge c_2$, reject $H_0$.

```{r, echo = FALSE}
print(p4)
```

### p-value {.fragile .allowframebreaks}

$\pval$ is the \textcolor{violet}{probability which generates the data}, \textcolor{blue}{or something else which is equal} \textcolor{red}{or rarer}.

$\pval$ is the probability of being $H_A$. 

\begin{center}
$t = `r round(t, 2)`$
\end{center}

```{r, echo = FALSE}
p_val <- pt(-abs(t), df)
p5 <- p2 + geom_vline(xintercept = -t, linetype = "dashed", 
                      col = "blue") +
  annotate("text", x = -t, y = 0.35, label = c("-t"), 
           hjust = -0.5, col = c("blue")) +
  geom_ribbon(data = subset(dat.st, x >= rx[1] & x <= -abs(t)),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA,
              alpha = 0.5) +
  geom_ribbon(data = subset(dat.st, x >= abs(t) & x <= rx[2]),
              aes(ymax = y), ymin = 0, fill = "red", colour = NA,
              alpha = 0.5)
print(p5)
```

\newpage

```{r}
p_value <- 2 * pt(-abs(t), df)
```
\begin{center}
$\pval = p_1 + p_2 = `r sprintf("%.3f + %.3f = %.3f", p_val, p_val, p_value)`$
\end{center}

```{r, echo = FALSE}
p6 <- p5 +
  annotate("segment", x = c(-3.5, 3.5), y = 0.1, 
             xend = c(-1.5, 1.5), yend = 0.05,
             arrow = arrow(angle = 20, length = unit(2, "mm"), 
                           type = "closed"), col = "brown") +
    annotate("label", x = c(-3.5, 3.5), y = 0.1,
           label = TeX(sprintf("$p_%d = %.3f$", 1:2, p_val),
                       output = "character"), parse = TRUE, col = "brown")
print(p6)
```

### Method-III (Decide tests result) {.fragile}

You can use $\pval$ to decide the test result. 

- If $\pval > \alpha$, **fail to reject** $H_0$.
- If $\pval \le \alpha$, reject $H_0$.

\begin{center}\rule{0.5\linewidth}{0.5000000000pt}\end{center}

As we **failed to reject** $H_0$ by three methods, we can conclude that our observations ($\bar{x}=`r round(mean(x), 2)`$) come from a normal distribution which $\mu = `r mu`$ with $`r 100 * conf.level`\%$ confidence. The means are not significantly different. Also, confidence interval is $CI = (`r sprintf("%.2f", ci)`)$.

## t-test function

### t.test 'equality' for mean {.fragile}

The steps above can be easly done by `t.test` funcion in R.

\begin{center}
$`r sprintf("t = %.3f, df = %d, \\pval = %.3f", t, df, p_value)`$

$CI = (`r sprintf("%.2f", ci)`)$
\end{center}

```{r}
t.test(x, mu = mu, conf.level = conf.level)
```

### t.test 'less than' for mean {.fragile}

1. $H_{0}$: $\bar{x}$ is equal or greater than $\mu$. ($\bar{x} \ge \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is less than $\mu$. ($\bar{x} > \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "less", mu = mu, conf.level = conf.level)
```

### t.test 'greater than' for mean {.fragile}

1. $H_{0}$: $\bar{x}$ is equal or less than $\mu$. ($\bar{x} \le \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is greater than $\mu$. ($\bar{x} > \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "greater", mu = mu, conf.level = conf.level)
```

## Which \ha{} should I use?
### Which \ha{} should I use?{.fragile}

\begin{exampleblock}{Question}

Which does it make sense? 

\begin{itemize}
\item Equality
\item Less
\item Greater
\end{itemize}

\end{exampleblock}

## Another t-test Example

### Another t-test Example {.fragile .allowframebreaks}

What if $\mu = 37$? Be aware confidence interval.

- $H_{0}$: $\bar{x}$ is equal to $\mu$. ($\bar{x} = \mu$), ($\pval > \alpha$).
- $H_{A}$: $\bar{x}$ is not equal to $\mu$. ($\bar{x} \ne \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, mu = 37, conf.level = conf.level)
```

\newpage

1. $H_{0}$: $\bar{x}$ is equal or greater than $\mu$. ($\bar{x} \ge \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is less than $\mu$. ($\bar{x} > \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "less", mu = 37, conf.level = conf.level)
```

\newpage

1. $H_{0}$: $\bar{x}$ is equal or less than $\mu$. ($\bar{x} \le \mu$), ($\pval > \alpha$).
2. $H_{A}$: $\bar{x}$ is greater than $\mu$. ($\bar{x} > \mu$), ($\pval \le \alpha$)

```{r}
t.test(x, alternative = "greater", mu = 37, conf.level = conf.level)
```

## Meaning of p-value

### What is the exact meaning of p-value? {.fragile .allowframebreaks}

Even if $H_0$ is \textcolor{red}{$true$}, there is still $5\%$ chance to reject $H_0$.
```{r}
try <- 1000
pvals <- sapply(1:try, function(i) {
  set.seed(i)
  t.test(rnorm(n))$p.value
})
per_of_type1 <- 100 * sum(pvals < 0.05)/try
```

```{r, echo = FALSE, fig.height = 1.5}
cat(sprintf("Percent of Type I Error : %1.f%%", per_of_type1))
breaks <- seq(0, 1, 0.05)
a <- c(1, rep(NA, length(breaks) - 2))
per <- function(x) {
  c(percent(x[1]/length(pvals)), rep("", length(x) - 1))
}
ggplot(data.frame(x = pvals), aes(x)) +
  geom_histogram(breaks = breaks, col = "black", 
                 fill = c("red", rep(NA, length(breaks) - 2)), alpha = 0.5) +
  stat_bin(breaks = breaks, aes(label = per(..count..)), 
           vjust = -0.5, geom = "text", na.rm = TRUE) +
  labs(x = "p-values")

# power.t.test(n, mu - mean(x), sd(x), alpha, type = "one.sample")
```

\newpage

