---
title: "Hypothesis Testing"
subtitle: "(t-test)"
abstract: "This is a presentation about Hypothesis Testing"
keywords: "hypothesis testing, t-test, R"
titlegraphic: "../images/itu.eps"
titlegraphic-width: 0.15
author:
  - \href{mailto:sezenismail@gmail.com}{İ.SEZEN}
  - \href{mailto:alper.unal@gmail.com}{A.ÜNAL}
institute:
  - Eurasia Institute of Earth Sciences
institute-short: İTÜ
date: \today
fontsize: 10pt
navigation: horizontal
graphics: yes
graphicspath:
  - \graphicspath{{hypo_files/figure-beamer/}}
beamerconfig:
    blocks: # default rounded, no shadow. Do not recommend shadows.
      shadow:
section-titles: # Hide section title pages
colorlinks: yes
linkcolor: "crimson"
toc: yes
handout: true
toc-title: Summary
toc-depth: 3
justifying: yes
output:
  beamer_presentation:
    template: ../template/default-2.9.2.1.beamer
    keep_tex: yes
    slide_level: 3
    fig_caption: yes
    theme: CambridgeUS
    colortheme: euresia
    highlight: tango
    pandoc_args: ["--pdf-engine-opt=-shell-escape", "--listings", "-H", "../template/listings-setup.tex"]
    # pandoc_args: ["--highlight-style", "my.theme"]
    includes:
      in_header: ../template/before_body.tex
      after_body: ../template/after_body.tex
---

```{r setup, include=FALSE}
source("../setup.R")
```

# \hypt
## What is \hypt?
### What is \hypt?

\hyp\ is,

- a supposition or proposed explanation made on the basis of limited evidence 
  as a starting point for further investigation.
- a testable statement about the relationship **between two or more variables**.
- a proposed explanation for some observed phenomenon^[[1]]^.
- an assumption about a population parameter. This assumption may or may not be 
  true (For instance, a population mean is equal to $10$)^[[2]]^.

[1]: https://whatis.techtarget.com/definition/hypothesis
[2]: https://stattrek.com/statistics/dictionary.aspx?definition=statistical_hypothesis

### What is \hypt?

\hyp\ is,

- a statistical method that is used in making statistical decisions using 
  experimental data^[[3]]^.
- basically an assumption that we make about the population parameter^[[3]]^.
- is a method of [statistical inference](https://en.wikipedia.org/wiki/Statistical_inference)^[[4]]^.
- The theory, methods, and practice of testing a \hyp\ by comparing it 
  with the **\hn{}**.

[3]: https://www.statisticssolutions.com/hypothesis-testing/
[4]: https://en.wikipedia.org/wiki/Statistical_hypothesis_testing

### Usage Cases

- A _sample_ is compared against a population parameter.
- Two statistical data sets are compared^[[3]]^.
- a data set obtained by sampling is compared against a synthetic data set 
  from an idealized model^[[3]]^.
  
### Steps of \hyp\ test

1. Identify a _\ts{}_ that is appropriate to the data and question.
2. Define a **\hn{}**, $H_{0}$.
3. Define an **\ha{}**, $H_{A}$.
4. Obtain the _null distribution_, which is simply the sampling distribution
  for the test statistic, if the _\hn{}_ is _true_.
5. Compare the observed _\ts{}_ to the _null distribution_.

### Example #1 {.fragile}

```{r code1, echo=TRUE, caption = "Hypothesis Testing"}
# utility functions utility functions utility functions
readinteger <- function(...) {
  a <- data.frame(matrix(1:20, ncol = 12, nrow = 20))
  ncol(a[,12:23])
  z <- median(a)
  x <- 2^35; integer(c(4321,3:56)) -> y; c <- NA
  a <<- x == y & x != y & x <= y | x >= y || x < y && x > y
  b <- !x + y - 1.5 * 12 / 0.3 %% 2 %/% 12 %*% 5 %in% a %o% x
  s <- "^[0-9]+$?"
  n <- !rnorm(x)
  plot(x, y, type = "l")
  if (length(n) == x) print(s)
  warning("message warning")
  stopifnot(1 == 1, all.equal(pi, 3.14159265), 1 < 2, 
            LETTERS, month.abb, airquality) # all TRUE
  return(as.integer(n))
}
```

## Sub-Section 2
### Frame Title 1
justify This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

# Section 2
## Sub-Section 1
### Frame Title 1
justify This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

